### 2019-11-08 latex/vhcs.tex
Finished VHC constrained dynamics
Next is hamiltonian

>>>>

By how \(C(q,\dot{q})\) is defined, 
one can collect the product \(B^\perp C \dot{q}\) in the following way.
First, combine the Christoffel symbols  into a matrix
\(Q_i : \mathcal{Q} \rightarrow \mathbb{R}^{n \times n}\)
defined element-wise by
\begin{equation*}
   [Q_i]_{j,k}(q) = \frac{1}{2}\left(\partial_{q_k} [D]_{i,j}(q)
      + \partial_{q_j} [D]_{i,k}(q) 
      + \partial_{q_i} [D]_{k,j}(q)\right)
\end{equation*}

Then \(B^\perp C \dot{q}\) becomes
\begin{equation*}
   \sum \limits_{i=1}^n B_i^\perp \dot{q}^T Q_i
\end{equation*}

To simplify the notation, define 
\begin{align*}
   \Psi_1(s) &:= -\frac{B^\perp \nabla P}{B^\perp D \sigma'} \\
   \Psi_2(s) &:= -\frac{B^\perp D \sigma" + \sum\limits_{i=1}^n B_i^\perp
   {\sigma'}^T Q_i \sigma'}{B^\perp D \sigma'} 
\end{align*}

Hence, the constrained dynamics for any regular VHC of order \((n-1)\) are
provided entirely by the single, second-order system 
(\ref{eqn:lagrangian_constrained_dynamics}).
\begin{equation}\label{eqn:lagrangian_constrained_dynamics}
   \ddot{s} = \Psi_1(s) + \Psi_2(s)\dot{s}^2
\end{equation}
